---
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(lattice); library(ggplot2); library(caret); library(randomForest); library(rpart); library(rpart.plot);
set.seed(20000)
```

#Data Processing

```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"   
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" 
```

#Data Cleaning

```{r}
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))  
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```

#Remove variables with near zero variance
```{r}
training<-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]
```
#Remove columns that are not predictors, which are the the seven first columns
```{r}
training   <-training[,-c(1:7)]
testing <-testing[,-c(1:7)]
```

#Cross-validation
In order to get out-of-sample errors, split the training data in training (75%) and testing (25%) data) subsets:
```{r, echo=FALSE}
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)    
NEOTraining <- training[inTrain, ]
NEOTesting <- training[-inTrain, ]  
dim(NEOTraining)
dim(NEOTesting) 

plot(NEOTraining$classe, col="yellow", main = "Levels of variable classe within the NEOTraining data set", xlab = "classe", ylab = "Frequency")
```

#Prediction

#DECISION TREE
#Fit model on Neo-Training data
```{r, echo=FALSE}
fitDT <- rpart(classe ~ ., data=NEOTraining, method="class")
```

#Use model to predict class in validation set (NEOTesting)
```{r, echo=FALSE}
predictionDT <- predict(fitDT, NEOTesting, type = "class")

rpart.plot(fitDT, main="Classification Tree", extra=102, under=TRUE, faclen=0)
```

#Estimate the errors of the prediction algorithm in the Decision Tree model
```{r, echo=FALSE}
confusionMatrix(predictionDT, NEOTesting$classe)
```

#RANDOM FOREST
#Fit model on NEOTraining data
```{r, echo=FALSE}
fitRF <- randomForest(classe ~ ., data=NEOTraining, method="class")
```

#Use model to predict class in validation set (NEOTesting)
```{r, echo=FALSE}
predictionRF <- predict(fitRF, NEOTesting, type = "class")
```

#Estimate the errors of the prediction algorithm in the Random Forest
```{r, echo=FALSE}
confusionMatrix(NEOTesting$classe, predictionRF)
```

#TEST THE MODEL TO PREDICT 20 DIFFERENT TEST CASES
#Perform prediction
```{r, echo=FALSE}
predictSubmission <- predict(fitRF, testing, type="class")
predictSubmission
```